prepare map backgrounds

### 2023-08-03 analysis/supplementary_materials/map backgrounds.Rmd

```{r}
africa_outline <- sf::st_read(here::here("analysis", "data", "raw_data", "Africa_outline.shp")) %>% sf::st_crop(xmax = 54, xmin = -20, ymax = 40, ymin = -36)
```

@@

```{r subset adm0}
mali_adm0 <- ne_countries(scale = "medium", returnclass = "sf") %>% filter(admin == "Mali")
```

@@

```{r plot adm0}
ggplot()+
  geom_sf(data = mali_adm0)+
  geom_sf_text(data = mali_adm0, aes(label = admin))

ggplot(africa_outline)+
  geom_sf()
```

@@

```{r}
rivers <- ne_download(scale = 50, type = "rivers_lake_centerlines", category = "physical", returnclass = "sf") %>% sf::st_intersection(mali_adm0) %>% filter (name == "Niger")

@@

```{r plot hydro}
ggplot()+
  geom_sf(data = mali_adm0)+
  geom_sf_text(data = mali_adm0, aes(label = admin))+
  geom_sf(data = rivers, fill = '#9ecae1')+
  geom_sf(data = osm_lakes.sf)
```

@@

#Mosaic/merge srtm tiles
srtmmosaic <- mosaic(srtm1, srtm2, srtm3, srtm4, srtm5, srtm6, srtm7, srtm8, srtm9, srtm10, fun=mean)
#crop to Mali
dem1 <- crop(srtmmosaic, extent(mali_sp))
mali_srtm <- mask(dem1, mali_sp)
#downsample
mali_srtm_a <- terra::aggregate(mali_srtm, fact = 10)
#convert to df
dem.df <- rasterToPoints(mali_srtm_a)
dem.df <- data.frame(dem.df)
colnames(dem.df) = c("lon", "lat", "alt")
```

@@

```{r}
# extract slope and aspect
slope.raster <- terrain(mali_srtm_a, opt='slope')
aspect.raster <- terrain(mali_srtm_a, opt='aspect')
# calculate hill shade
hill.raster <- hillShade(slope.raster, aspect.raster, 40, 270)
# convert to df
hill.m <- rasterToPoints(hill.raster)
hill.df <-  data.frame(hill.m)
colnames(hill.df) <- c("lon", "lat", "hill")
```

```{r}
ggplot()+
  #geom_sf(data = mali_adm0)+
  geom_raster(data = hill.df, aes(lon, lat, fill = hill)) +
  scale_fill_gradientn(colours = grey.colors(100)) +
  geom_raster(data = dem.df, aes(lon, lat, fill = alt), alpha = 0.8)+
  scale_fill_gradientn(colours = terrain.colors(100))+
  geom_sf_text(data = mali_adm0, aes(label = admin))+
  geom_sf(data = rivers, colour = "blue")+
  theme_minimal()

>>>>

```{r}
africa_outline <- sf::st_read(here::here("analysis", "data",
                                         "raw_data", "Africa_outline.shp")) |>
  sf::st_crop(xmax = 54, xmin = -20, ymax = 40, ymin = -36)
```

@@

```{r subset adm0}
mali_adm0 <- ne_countries(scale = "medium", returnclass = "sf") |>
  filter(admin == "Mali")
```

@@

```{r plot adm0}
ggplot() +
  geom_sf(data = mali_adm0) +
  geom_sf_text(data = mali_adm0, aes(label = admin))

ggplot(africa_outline) +
  geom_sf()
```

@@

```{r}
rivers <- ne_download(scale = 50, type = "rivers_lake_centerlines",
                      category = "physical", returnclass = "sf") |>
  sf::st_intersection(mali_adm0) |> filter(name == "Niger")

@@

```{r plot hydro}
ggplot() +
  geom_sf(data = mali_adm0) +
  geom_sf_text(data = mali_adm0, aes(label = admin)) +
  geom_sf(data = rivers, fill = '#9ecae1') +
  geom_sf(data = osm_lakes.sf)
```

@@

#Mosaic/merge srtm tiles
srtmmosaic <- mosaic(srtm1, srtm2, srtm3, srtm4, srtm5, srtm6, srtm7, srtm8, srtm9, srtm10, fun = mean)
#crop to Mali
dem1 <- crop(srtmmosaic, extent(mali_sp))
mali_srtm <- mask(dem1, mali_sp)
#downsample
mali_srtm_a <- terra::aggregate(mali_srtm, fact = 10)
#convert to df
dem.df <- rasterToPoints(mali_srtm_a)
dem.df <- data.frame(dem.df)
colnames(dem.df) <- c("lon", "lat", "alt")
```

@@

```{r}
# extract slope and aspect
slope.raster <- terrain(mali_srtm_a, opt = 'slope')
aspect.raster <- terrain(mali_srtm_a, opt = 'aspect')
# calculate hill shade
hill.raster <- hillShade(slope.raster, aspect.raster, 40, 270)
# convert to df
hill.m <- rasterToPoints(hill.raster)
hill.df <-  data.frame(hill.m)
colnames(hill.df) <- c("lon", "lat", "hill")
```

```{r}
ggplot() +
  geom_raster(data = hill.df, aes(lon, lat, fill = hill)) +
  scale_fill_gradientn(colours = grey.colors(100)) +
  geom_raster(data = dem.df, aes(lon, lat, fill = alt), alpha = 0.8) +
  scale_fill_gradientn(colours = terrain.colors(100)) +
  geom_sf_text(data = mali_adm0, aes(label = admin)) +
  geom_sf(data = rivers, colour = "blue") +
  theme_minimal()
